---
title       : Predict a Child's Height 
subtitle    : Based on data from the Galton dataset
author      : Emily Jones
job         : Coursera Developing Data Products Course Project
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Web Application to Predict a Child's Height based on their Parent's Height

* The web app is found at http://mononajones.shinyapps.io/Galton/

* It uses the dataset GaltonFamilies in the HistData package. 

This dataset is based on the work done by Francis Galton in the 19th century

The help from the HistData package states that data set lists the individual observations for 934 children in 205 families on which Galton (1886) based his cross-tabulation.

--- .class #id 

## Functionality
<p style='font-size:smaller'>
This application allows one to select the height for each parent. It uses a linear regression model to predict a child's height based on the these heights. It then returns predictions for the height of sons and daughters.
The following is the code used by the application:
</p>
<p style='font-family:"Lucida Console"; font-size:small'>
 fit<-lm(childHeight~mother+father+gender, data=GaltonFamilies)
 <br/>
 predict.lm(fit, data.frame(mother=m,father=f,gender=g), interval="prediction")
</p>
<p style='font-size:75%'>
where m is the height of the mother in inches, f is the height of the father in inches and g is female for daughters and male for sons.
</p>
```{r fit, echo=FALSE}
    require("HistData")
    # create the fit 
    fit<- lm(childHeight~mother+father+gender, data=GaltonFamilies)
    m=64
    f=72
    # predict for daughters
    d<- predict.lm(fit, data.frame(mother=m,father=f,gender="female"), interval="prediction")
    # predict for sons
    s<-predict.lm(fit, data.frame(mother=m,father=f,gender="male"), interval="prediction")

```
<p style='font-size:75%'>
Given a mother who is 64 inches tall and a father who is 72 inches tall, this predicts a daughter's height as follows:
</p>
```{r daughter, echo=FALSE}
    m=64
    f=72
    # predict for daughters
   predict.lm(fit, data.frame(mother=m,father=f,gender="female"), interval="prediction")
```
<p style='font-size:75%'>
This predicts a son's height as follows: 
</p>
```{r son, echo=FALSE}
    m=64
    f=72
    # predict for daughters
   predict.lm(fit, data.frame(mother=m,father=f,gender="male"), interval="prediction")
```

--- .class #id 
## Example Output
### Screen Capture of the application


  <img src = 'GaltonResults.PNG' style='width:750 ;height: 350' ></img>

--- .class #id 

## Source and References

### Source:
<p>Galton's notebook, http://www.medicine.mcgill.ca/epidemiology/hanley/galton/notebook/, transcribed by Beverley Shipley in 2001.
</p>
### References
References for the Galton dataset as provided by the DataHist package.
<p>Galton, F. (1886). Regression Towards Mediocrity in Hereditary Stature Journal of the Anthropological Institute, 15, 246-263
</p>
<p>
Hanley, J. A. (2004). "Transmuting" Women into Men: Galton's Family Data on Human Stature. The American Statistician, 58, 237-243. See: 
<br/> http://www.medicine.mcgill.ca/epidemiology/hanley/galton/ for source materials.
</p>
<p>
Keen, K. J. (2010). Graphics for Statistics and Data Analysis with R, Boca Raton: CRC Press,
<br/>http://www.unbc.ca/keen/graphics-for-statistics-and-data-analysis-with-r.
</p>

